---
title: "Protests against Trump, January 1st, 2025 to May 31st, 2025"
author: "Philippe Joly"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

This project uses data from the Crowd Counting Consortium (CCC) to analyze protests against Trump in 2025 (from January 1st, 2025 to October 31st, 2025).

```{r load-libs, include=FALSE}
Sys.setlocale("LC_TIME", "C")
library(dplyr)
library(ggplot2)
library(ggrepel)
library(lubridate)
library(MASS)
library(marginaleffects)
library(modelsummary)
```

```{r load-data}
ccc_events <- readRDS("data/ccc_events.rds")

ccc_states <- readRDS("data/ccc_states.rds") |> 
  filter(!(state %in% c("DC", "PR")))
```

## Trends over Time

### Figure 1

```{r time-trend-events}
data_plot <- ccc_events |> 
  mutate(month = month(date)) |> 
  group_by(month, date) |> 
  summarize(n = n()) |> 
  mutate(label = case_match(date,
                            as.Date("2025-10-18") ~ "No Kings 2.0 (Oct 18) → ",
                            as.Date("2025-04-05") ~ "Hands Off (Apr 5)→ ")) |> 
  arrange(desc(n)) |> 
  group_by(month) |> 
  mutate(avg_n = sum(n) / n())

data_plot |> 
  ggplot(aes(x = date, y = n, label = label)) +
  geom_col() +
  geom_line(aes(y = avg_n), color = "red") +
  geom_text(hjust = 1, vjust = 0.5) +
  labs(title = "Number of Events per Day",
       x = "Date",
       y = "Number of events per day",
       caption = "Note: The red line shows the average daily number of events per month.") + 
  theme_minimal() +
  theme(legend.title=element_blank()) 
```

### Figure 2

```{r time-trend-part}
data_plot <- ccc_events |> 
  mutate(month = month(date)) |> 
  group_by(month, date) |> 
  summarize(n_part = sum(size_mean, na.rm = TRUE) / 1000) |> 
  mutate(label = case_match(date,
                            as.Date("2025-10-18") ~ "No Kings 2.0 (Oct 18) → ",
                            as.Date("2025-04-05") ~ "Hands Off (Apr 5)→ ")) |> 
  arrange(desc(n_part)) |> 
  group_by(month) |> 
  mutate(avg_n_part = sum(n_part) / n())

data_plot |> 
  ggplot(aes(x = date, y = n_part, label = label)) +
  geom_col() +
  geom_line(aes(y = avg_n_part), color = "red") +
  geom_text(hjust = 1, vjust = 0.5) +
  labs(title = "Number of Participants per Day",
       subtitle = "In thousands",
       x = "Date",
       y = "Number of participants per day\n(in thousands)",
       caption = "Note: The red line shows the average daily number of participants per month (in thousands).") + 
  theme_minimal() +
  theme(legend.title=element_blank()) 
```

### Figure 3

```{r time-trend-violence}
data_plot <- ccc_events |> 
  mutate(month = month(date),
         violence = if_else(property_damage_any == 1 | police_casualties_any == 1,
                           "Violent",
                           "Nonviolent"),
         violent = factor(violence)) |> 
  group_by(month, violence) |> 
  summarize(n = n()) |> 
  mutate(share = (n/sum(n)) * 100) |> 
  filter(violence == "Violent")

data_plot |> 
  mutate(month = factor(month, levels = 1:12, labels = month.abb),
         label = sprintf("%.2f", share)) |> 
  ggplot(aes(x = month, y = share, label = label)) +
  geom_col() +
  geom_text(vjust = -0.5) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(title = "Percentage of Violent Protest Events", 
       subtitle = "By month",
       x = "Month",
       y = "Percentage of violent protests events") +
  theme_minimal()
```

## Geography of the Protests: Bivariate Relationships

### Figure 4

```{r trump-support}
ccc_states|> 
  filter(!is.na(pct_trump), !is.na(n_part_pop), state != "DC") |> 
  ggplot(aes(x = pct_trump, y = n_part_pop)) +
  geom_point() +
  geom_smooth(
    method = "glm",
    method.args = list(family = "poisson"),
    formula = "y ~ x"
  ) +
  geom_text_repel(aes(label = state)) +
  labs(title = "Number of Participants by Support for Trump in the 2024 Presidential Election", 
       subtitle = "By state, per thousand inhabitants",
       x = "Support for Trump (%), 2024 presidential election",
       y = "Number of participants per thousand inhabitants") +
  theme_minimal()
```

### Figure 5

```{r income}
ccc_states|> 
  filter(!is.na(income), !is.na(n_part_pop), state != "DC") |> 
  ggplot(aes(x = income, y = n_part_pop)) +
  geom_point() +
  geom_smooth(
    method = "glm",
    method.args = list(family = "poisson"),
    formula = "y ~ x"
  ) +
  geom_text_repel(aes(label = state)) +
  labs(title = "Number of Participants by Median Annual Household Income", 
       subtitle = "By state, per thousand inhabitants",
       x = "Median annual household income (in 1,000$)",
       y = "Number of participants per thousand inhabitants") +
  theme_minimal()
```

### Figure 6

```{r black}
ccc_states|> 
  filter(!is.na(black), !is.na(n_part_pop), state != "DC") |> 
  ggplot(aes(x = black, y = n_part_pop)) +
  geom_point() +
  geom_smooth(
    method = "glm",
    method.args = list(family = "poisson"),
    formula = "y ~ x"
  ) +
  geom_text_repel(aes(label = state)) +
  labs(title = "Number of Participants by Percentage of Black Residents", 
       subtitle = "By state, per thousand inhabitants",
       x = "Percentage of Black residents",
       y = "Number of participants per thousand inhabitants") +
  theme_minimal()
```

### Figure 7

```{r hispanic}
ccc_states|> 
  filter(!is.na(hispanic), !is.na(n_part_pop), state != "DC") |> 
  ggplot(aes(x = hispanic, y = n_part_pop)) +
  geom_point() +
  geom_smooth(
    method = "glm",
    method.args = list(family = "poisson"),
    formula = "y ~ x"
  ) +
  geom_text_repel(aes(label = state)) +
  labs(title = "Number of Participants by Percentage of Hispanic Residents", 
       subtitle = "By state, per thousand inhabitants",
       x = "Percentage of Hispanic Residents",
       y = "Number of participants per thousand inhabitants") +
  theme_minimal()
```

### Figure 8

```{r non-citizens}
ccc_states|> 
  filter(!is.na(pct_non_citizens), !is.na(n_part_pop), state != "DC") |> 
  ggplot(aes(x = pct_non_citizens, y = n_part_pop)) +
  geom_point() +
  geom_smooth(
    method = "glm",
    method.args = list(family = "poisson"),
    formula = "y ~ x"
  ) +
  geom_text_repel(aes(label = state)) +
  labs(title = "Number of Participants by Percentage of Noncitizen Residents", 
       subtitle = "By state, per thousand inhabitants",
       x = "Percentage of noncitizen residents",
       y = "Number of participants per thousand inhabitants") +
  theme_minimal()
```

### Figure 9

```{r gov-party}
ccc_states|> 
  filter(!is.na(gov_party), !is.na(n_part_pop), state != "DC") |> 
  ggplot(aes(x = gov_party, y = n_part_pop)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "centerwhole", size = 0.5) +
  geom_text_repel(aes(label = state)) +
  labs(title = "Number of Participants by Party of the Governor", 
       subtitle = "By state, per thousand inhabitants",
       x = "Governor party",
       y = "Number of participants per thousand inhabitants",
       caption = "Note: The center of the boxplot is the median; the lower and upper hinges represent the 25th and 75th percentiles.") +
  theme_minimal()
```

## Geography of the Protests: Multivariate Analysis

### Figure 10

Results of a count model. Coefficients below zero indicate a negative effect. 

```{r multivariate}
model <- glm.nb(
  n_part ~ 
    pct_trump + 
    income + 
    black +
    hispanic + 
    pct_non_citizens + 
    gov_party +
    offset(log(pop)),
  data = ccc_states
)

mfx <- avg_comparisons(model)

cm <- c('pct_trump'    = 'Trump support (%), 2024 election',
        'income' = "Annual median income",
        'black' = "Black residents (%)",
        'hispanic' = "Hispanic residents (%)",
        'pct_non_citizens' = "Noncitizen Residents (%)",
        'gov_party' = "Republican governor")

modelplot(mfx, coef_map = cm) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Effects of different variables",
       subtitle = "All things being equal",
       x = "Expected effect of one-unit change\n(with 95% confidence intervals)",
       caption = "Note: When the confidence interval overlaps with zero, the effect is not statistically significant.")
```
