---
title: "Protests against Trump, January 1st, 2025 to May 31st, 2025"
author: "Philippe Joly"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

This project uses data from the Crowd Counting Consortium (CCC) to analyze protests against Trump in 2025 (from January 1st, 2025 to October 31st, 2025).

```{r load-libs}
library(dplyr)
library(ggplot2)
library(ggrepel)
library(lubridate)
library(modelsummary)
```

```{r load-data}
ccc_events <- readRDS("data/ccc_events.rds")

ccc_states <- readRDS("data/ccc_states.rds") |> 
  filter(!(state %in% c("DC", "PR")))
```

## Protests over Time

```{r time-trend}
data_plot <- ccc_events |> 
  mutate(month = month(date)) |> 
  group_by(month, date) |> 
  summarize(n_part = sum(size_mean, na.rm = TRUE) / 1000) |> 
  mutate(label = case_match(date,
                            as.Date("2025-10-18") ~ "No Kings 2.0 (Oct 18) → ",
                            as.Date("2025-04-05") ~ "Hands Off (Apr 5)→ ")) |> 
  arrange(desc(n_part)) |> 
  group_by(month) |> 
  mutate(avg_n_part = sum(n_part) / n())

data_plot |> 
  ggplot(aes(x = date, y = n_part, label = label)) +
  geom_col() +
  geom_line(aes(y = avg_n_part, color = "Average daily number\nof participants per month\n(in thousands)")) +
  geom_text(hjust = 1, vjust = 0.5) +
  labs(x = "Date",
       y = "Number of participants per day\n(in thousands)") + 
  theme_minimal() +
  theme(legend.title=element_blank()) 
```



## Bivariate Relationships

### Support for Trump (%), 2024 presidential election

```{r trump-support}
ccc_states|> 
  filter(!is.na(pct_trump), !is.na(n_part_pop), state != "DC") |> 
  ggplot(aes(x = pct_trump, y = n_part_pop)) +
  geom_point() +
  geom_smooth(
    method = "glm",
    method.args = list(family = "poisson"),
    formula = "y ~ x"
  ) +
  geom_text_repel(aes(label = state)) +
  labs(x = "Support for Trump (%), 2024 presidential election",
       y = "Number of participants per thousand inhabitants") +
  theme_minimal()
```

### Median Annual Household Income

```{r income}
ccc_states|> 
  filter(!is.na(income), !is.na(n_part_pop), state != "DC") |> 
  ggplot(aes(x = income, y = n_part_pop)) +
  geom_point() +
  geom_smooth(
    method = "glm",
    method.args = list(family = "poisson"),
    formula = "y ~ x"
  ) +
  geom_text_repel(aes(label = state)) +
  labs(x = "Median annual household income (in 1,000$)",
       y = "Number of participants per thousand inhabitants") +
  theme_minimal()
```

### Percentage of Black Residents

```{r black}
ccc_states|> 
  filter(!is.na(black), !is.na(n_part_pop), state != "DC") |> 
  ggplot(aes(x = black, y = n_part_pop)) +
  geom_point() +
  geom_smooth(
    method = "glm",
    method.args = list(family = "poisson"),
    formula = "y ~ x"
  ) +
  geom_text_repel(aes(label = state)) +
  labs(x = "Percentage of Black Residents",
       y = "Number of participants per thousand inhabitants") +
  theme_minimal()
```

### Percentage of Hispanic Residents

```{r hispanic}
ccc_states|> 
  filter(!is.na(hispanic), !is.na(n_part_pop), state != "DC") |> 
  ggplot(aes(x = hispanic, y = n_part_pop)) +
  geom_point() +
  geom_smooth(
    method = "glm",
    method.args = list(family = "poisson"),
    formula = "y ~ x"
  ) +
  geom_text_repel(aes(label = state)) +
  labs(x = "Percentage of Hispanic Residents",
       y = "Number of participants per thousand inhabitants") +
  theme_minimal()
```

## Percentage of Noncitizen Residents

```{r non-citizens}
ccc_states|> 
  filter(!is.na(pct_non_citizens), !is.na(n_part_pop), state != "DC") |> 
  ggplot(aes(x = pct_non_citizens, y = n_part_pop)) +
  geom_point() +
  geom_smooth(
    method = "glm",
    method.args = list(family = "poisson"),
    formula = "y ~ x"
  ) +
  geom_text_repel(aes(label = state)) +
  labs(x = "Noncitizen Residents (%)",
       y = "Number of participants per thousand inhabitants") +
  theme_minimal()
```

## Governor Party

```{r gov-party}
ccc_states|> 
  filter(!is.na(gov_party), !is.na(n_part_pop), state != "DC") |> 
  ggplot(aes(x = gov_party, y = n_part_pop)) +
  geom_boxplot() +
  geom_point() +
  geom_text_repel(aes(label = state)) +
  labs(x = "Governor party",
       y = "Number of participants per thousand inhabitants") +
  theme_minimal()
```

## Correlation Matrix

```{r cor}
ccc_states|> 
  select(n_part_pop, 
         pct_trump, 
         income, 
         black, 
         hispanic, 
         non_citizens = pct_non_citizens, 
         republican_gov = gov_party) |> 
  mutate(republican_gov = as.numeric(republican_gov)) |> 
  datasummary_correlation(coef_map = cm)

```

## Multivariate Analysis

Results of a count model. Coefficients below zero indicate a negative effect. 

```{r multivariate}
model <- glm(n_part ~ 
               pct_trump + 
               income + 
               black +
               hispanic + 
               pct_non_citizens + 
               gov_party +
               offset(log(pop)), 
             data = ccc_states, 
             family = poisson)

cm <- c('pct_trump'    = 'Trump support (%), 2024 election',
        'income' = "Annual median income",
        'black' = "Black residents (%)",
        'hispanic' = "Hispanic residents (%)",
        'pct_non_citizens' = "Noncitizen Residents (%)",
        'gov_partyRepublican' = "Republican governor")

modelplot(model, coef_map = cm) +
   geom_vline(xintercept = 0, color = "red", linetype = "dashed")
```
